The weight of the mountains: Constraints on tectonic stress, friction, and fluid pressure in the 2008 Wenchuan earthquake from estimates of topographic loading

### 2015-02-13 revisions/rebuttal.md

The marginal priors for
$T_{max}$ would be much more strongly biased in the first sampling approach,
while the marginals for
$T_{min}$ is less affected by the sampling strategy. Overall, we feel that our
approach is less biased, because it results in a uniform $T_{max}$, which is
the more important value.

On further reflection, it would be possible to pick a uniform $T_{min}$ and
then build $T_{max}$ around that, but then $T_{max}$ would be non-uniform. We
can think of no particular reason to favor this approach, since researchers are
generally more interested in estimating maximum principal tectonic stress than
minimum.

However, it is clear in Figure 8b, where $T'_{min}$ is scaled to $T'_{max}$,
that the *ratio* of $T_{min}$ to $T_{max}$ is constrained by the inversion;
the most likely value for $T_{min} \approx 0.3 T_{max}$, regardless of the
value for $T_{max}$. This point was mentioned in passing in section 5.1 (line
330) but could be discussed in more detail, particularly with reference to 
Figure 8.1. Therefore, we add to Section 5.2:

>It is also apparent in Figure 8b that regardless of the magnitude of
$T^\_{\mathrm{max}}$, $p_{J}(T^\prime_{\mathrm{min}} | D)$ is
consistently 0--0.6 $p_{J}(T^\prime_{\mathrm{max}} | D)$, with a mode
of about 0.3; this indicates that the relative magnitude of the
tectonic stresses has a substantial influence on the rake of the maximum
shear stress resolved on the fault.

### 2015-02-13 revisions/styron_hetland_wenchuan_topo_revised.tex
eric's changes

% document
\begin{abstract}
Though it is widely recognized that large mountain ranges produce significant
stresses in the Earth's crust, these stresses are not commonly quantified.
Nonetheless, near large mountains topography may affect fault activity by
changing the stress balance on the faults. In this work, we calculate the
stress field from topography in the Longmen Shan (Sichuan, China) and resolve
those stresses on several models of the faults that ruptured in the 2008
M$_\mathrm{w}$ 7.9 Wenchuan earthquake. We find that the topography results in
shear stresses up to 20 MPa and normal stresses up to 80 MPa on the faults,
with significant variability across the faults. Topographic stresses generally
load the fault in a normal and left-lateral shear sense, opposite to the
inferred coseismic slip sense, and thus inhibit the coseismic slip. We estimate
the accumulated tectonic stress needed to overcome topographic and lithostatic
stresses so that the direction of maximum shear on the faults is roughly
collinear with the inferred coseismic slip. We further estimate the static
friction and pore fluid pressure assuming that the fault was, on average, at
Mohr-Coulomb failure at the time of the Wenchuan earthquake. We use a Bayesian
inversion strategy, yielding posterior probability distributions for the
estimated parameters. We find most likely estimates of maximum tectonic
compressive stress near 0.6 $\rho g z$ and oriented $\sim$E-W, and minimum
tectonic stress near 0.2 $\rho g z$. Static friction on the fault is near 0.2,
and pore fluid pressure is between 0 and 0.4 of the total lithostatic
pressure.  \end{abstract}

@@

In areas of substantial relief, high terrain and steep slopes generate large
stresses in the crust beneath and adjacent to the high topography
\citep{jeffreys1924, coblentz1996}. Because of the irregularity of topography
in mountainous regions, the stresses produced by topography are also
heterogeneous, and may play a prominent role in local or regional deformation,
particularly if the region is tectonically active.  For example, shear and
normal stresses on a fault from topographic loading may push a particular fault
closer to or farther from failure, or reorient the
net shear stress direction on a fault. These effects may affect the
localization or deformational style in a region. Furthermore, heterogeneous
topographic stresses on a particular fault may affect the way earthquake
ruptures propagate across the fault plane. Despite this, the degree to which
topographic stresses affect faulting has received little direct study.

In this work, we investigate the effects of topographic stresses on the faults
that ruptured in the 2008 \emph{M}7.9 Wenchuan, China earthquake.  This
earthquake is an ideal candidate for this study because it occurred at the base
of the Longmen Shan, one of the largest and steepest escarpments on Earth
(Figure~\ref{fig:lms_map}), and has a well-studied coseismic slip distribution
characterized by significant along-strike variations in coseismic slip.
Additionally, because the earthquake occurred after $\sim$2000 years
of seismic quiescence, postseismic stresses in the lithosphere are likely to be
negligible, suggesting that the stress state on the faults at the time of
rupture can be approximated as the sum of topographic, lithostatic, and
accumulated tectonic stresses. We then use the Mohr-Coulomb failure criterion
to bracket the tectonic stress field, pore fluid pressure, and static friction
of the fault, so that these parameters are consistent with the toopographic
stresses and coseismic slip on the Wenchuan earthquake faults.

@@

Topographic stresses are only one component of the total stress field in the
crust \citep{molnar1988}. Coseismic slip in the Wenchaun earthquake is due to
the total accumulated stress on the faults, which also includes components from
lithostatic and tectonic stresses. (In the present study, we do not consider
stresses due to flexure \citep[e.g.,][] {luttrell2007} or buoyancy
\citep[e.g.,][]{luttrell2011}.) By quantifying both the topographic and
lithostatic stresses, we can use coseismic slip models to solve for the
tectonic stress assuming that (1) the total shear stress on the fault is in the
direction of coseismic slip \citep[e.g.,] []{angelier1994} and (2) the fault is
at Mohr-Coulomb failure everywhere that it slipped. If topographic stresses are
significant and produce shear in the direction of fault slip, then for given
values of static friction and pore fluid pressure, we can calculate the amount
of tectonic stress that can be added to the ambient stress field before the
faults should rupture; given limited acceptable ranges for friction and fluid
pressure, we are essentially able to place maximum constraints on tectonic
stress. Alternately, if topographic stresses work against coseismic slip, for
given friction and fluid pressures we can estimate the minimum magnitudes of
tectonic stresses necessary to overcome shear and frictional resistance to
slip. In a scenario with complex faulting and topography, it may be possible to
put bounds on both minimum and maximum magnitudes, in addition to the 
directions of tectonic
stresses. To account for the non-uniqueness of the solution, we use a Bayesian
Monte-Carlo methodology to estimate posterior probability density functions
(PDFs) of tectonic stresses, static fault friction, and pore fluid pressure.

@@

\begin{equation}
M(x, y, z) = G(x, y, z) * F(x, y) \; ,
\end{equation}
where $G(x,y,z)$ is a set of Green's functions for the six stress tensor
elements, and $F(x,y)$ is a topographic loading function, described
below. We assume compressive stresses are positive, and $x>0$ is east,
$y>$ is north, and $z>0$ is depth. \citet{liuzoback1992}
show that $M(x,y,z)$ can be decomposed into two components as

@@

\begin{equation}
M^C(x, y, z) = G_x^C(x,y,z) * F_{h, x}(x, y) + G_y^C(x,y,z) * F_{h, y}(x, y) \; ,
\end{equation}
where $G_i^C(x,y,z)$ are the Cerruti solutions for a horizontal point
source load in the $i$ direction on the halfspace surface (see Appendix
A1). The horizontal loading functions are given by

@@

Topographic stresses on the Wenchuan faults are calculated on point sets
representing the faults taken from coseismic slip models. We use six models,
those of \citet{shen2009}, \citet{feng2010}, \citet{zhang2011},
\citet{fielding2013} and two from \citet{qi2011}.
All of these models rely on geodetic data to some degree, but they do not all 
use the same data in their inversion (e.g., \citep{feng2010} uses a different
InSAR catalogue as the others). All use different inversion strategies, and
different fault geometries (the two models of \citep{qi2011} share a common
fault geometry but use different regularization in the inference of the slip
distribution). By using a suite of models in our calculations, we can infer
that results which are persistent in most or all of the models are more robust,
while other results specific to only one of the coseismic slip models can be
more confidently linked to specifics of the model geometry or inferred slip
distribution in those models.

@@

In our stress calculations, we discard
points above 851 m below sea level, as this is above the depth at which
we compute $G^C(x,y,z)$. The six stress tensor components calculated at
the regular grid points are linearly interpolated to points describing
the faults. Because the fault points are completely surrounded by the
grid nodes at which topographic stresses were calculated and those nodes
are spaced \textless{}1 km apart, the fault points cannot be more than a
few hundred meters from the nearest grid node, so a higher order
interpolation is not necessary. We then project the topographic stress
tensor to fault normal stress, $\sigma_n^M$, down-dip shear stress,
$\tau_d^M$ and strike-slip shear stress, $\tau_s^M$ at each point in
describing the fault geometry. In this work, the superscripts on stress
symbols denote the origin of the stresses.

@@

\begin{equation} 
\tau = \mu ( \sigma_n - \sigma_p ) \; ,
\label{eqn:amonton_raw} 
\end{equation}
where $\mu$ is the coefficient of static friction on the fault and
$\sigma_p$ is the pore fluid pressure \citep[e.g.,][]{sibson1985}, and
assuming that rock strength (cohesion) is negligible. We
describe the pore fluid pressure using a scalar, $0 \leq \phi \leq 1$,
which is the pore fluid pressure as a fraction of total pressure, and so
the failure criterion is

@@

We estimate the tectonic stress tensor field, $\mu$, and $\phi$
consistent with published coseismic slip models of the Wenchuan
earthquake using a Bayesian estimation, resulting in posterior probability 
density functions of the model parameters. 
We first estimate posteriors of the tectonic stress tensor field 
$T$ consistent with
the coseismic slip models, and then estimate $\mu$ and $\phi$ consistent
with Mohr-Coulomb failure. The nature of Bayesian estimation allows us
to quantify both the relative likelihoods of model parameters and the
tradeoffs between them.

@@

We invert topographic stresses and coseismic slip models for tectonics
stresses using Bayesian methods, and making the common `Wallace-Bott'
assumption (named after \citet{wallace1951} and \citet{bott1959}) 
that slip on the fault occurs in the general direction
of the maximum resolved shear stress on the fault \citep[e.g.,][]
{mckenzie1969, angelier1994}. We estimate the tectonic stresses in
light of the topographic stresses and slip distributions through the
relation

@@

Previous workers have demonstrated that by quantifying topographic
stress, other components in the Coulomb stress balance may be bracketed
\citep[e.g.,][]{cattin1997, lamb2006, luttrell2011}. Each of these studies
uses somewhat different approaches. Our approach is most similar to that
of \citet{luttrell2011}, although there are significant
differences: (1) We use the topographic stresss calculations proposed by
\citet{liuzoback1992}, whereas 
\citet{luttrell2011} only uses the vertical loading from topography,
equivalent to $M^B$ in equation (\ref{eqn:bous}). (2) We do not consider
bouyancy forces due to lateral variations in density, due for instance to
Moho variation, as done by \citet{luttrell2011}. In the
Longmen Shan region, Moho variation is small compared to the change in
Moho depths at the Andean plate boundary, and so the buoyancy terms
should be relatively small. (3) We consider a full range of tectonic
stresses, instead of simply calculating the minimum principal tectonic
stress and its orientation. (4) We consider the stress tensor at each
point due to topographic loading, lithostatic stress, and horizontal
tectonic stress, and use inferred coseismic slip rake as a constraint on
the allowable stresses, rather than inferring the earthquake stress drop
from the coseismic slip models, as done by 
\citet{luttrell2011}. (5) We use both normal and shear stresses to
constrain pore fluid pressure and friction. (6) We use a Bayesian
estimation, resulting in a PDF of tectonic stress, as well as $\mu$ and
$\phi$, rather than just solving for the minimum tectonic stress
required for faulting, as done by \citet{luttrell2011}.

@@

The maximum tectonic stress, $T_{\mathrm{max}}$, is consistently
oriented roughly E-W in our results. This orientation is oblique to the
Longmen Shan, which produces oblique (right-lateral and reverse sense)
shear on the Beichuan fault. $T_{\mathrm{min}}$ is $\sim$N-S
oriented, and is a small fraction of lithostatic pressure. This
stress configuration is in close agreement with pre-earthquake stress
orientation
measurements near the rupture zone (Figure \ref{fig:lms_stress_map}),
mostly from borehole breakout data from 2-5 km depth
\citep{heidbach2009}, which is the zone of maximum slip in the coseismic
slip models. It is somewhat discrepant with stress orientations
estimated at $\sim$800 m depth adjacent to the Beichuan fault
in the WFSD-1 borehole of the Wenchuan Earthquake Fault Scientific
Drilling Project several years after the 2008 earthquake \citep{cui2014},
which show $\sigma_{H{\mathrm{\max}}}$ to be more orthogonal to the fault
trace, suggesting that much of the right-lateral component of shear
stress was released during the earthquake. These results are similar to
the orientations of total stress obtained by 
\citet{medinaluna2013}, although they were unable to constrain the
magnitudes of stresses.

@@

The magnitudes of $T_{\mathrm{max}}$ and $T_{\mathrm{min}}$ are
dominantly constrained on the low end by our analysis, which is apparent
by the sharp decrease in the frequency of $p(T_{\mathrm{max}}|D)$ below
about $T^\prime_{\mathrm{max}} = 0.5$ (Figure~\ref{fig:joint_posteriors}). 
These results indicates that $T_{\mathrm{max}}$ of at
least $\sim$13.25 MPa km$^{-1}$ is necessary to overcome
topographic stresses resisting reverse and right-lateral slip on the
faults. We find that the highest likely ratio of strike-slip to dip-slip
shear along the Wenchuan earthquake faults is close to 1. This is
similar to the inferences of strain accumulation rates inferred from
squishy block modeling by \citet{loveless2011}, who
concluded that the rate of slip deficit accumulation in the thrust and
dextral senses were approximately equal. It should be noted that the
tectonic stresses we estimate here represent the accumulated stresses
prior to the Wenchuan earthquake, and the relation of these stresses to any
accumulated slip deficit needs to be through a model of strain
accumulation.

@@

CFR indicates whether a fault should fail under a given stress state:
CFR \textgreater{} 1 indicates failure, while a CFR \textless{} 1
indicates fault stability. We then calculate the CFR on each point in
the model of the Beichuan fault (594 points describe the fault model of
\citet{zhang2011}) based on the full stress field $S$ at
each point on the fault, for each of the 1000 samples of $T$, $\phi$ and
$\mu$ drawn randomly from the posteriors. We call this CFR$_f$. Then,
using the same $S$ and $\phi$, we calculate the CFR on an optimally
oriented fault with $\mu=0.6$ and no cohesion, which we call CFR$_o$.
The orientation of the optimally oriented fault is determined as being
the angle $\beta$ away from $\sigma^S_1$, where $\beta = (\tan^{-1} \mu) / 2$
and is in the $\sigma_1$--$\sigma_3$ plane \citep[e.g.,][]{sibson1985}.
Note this $\mu$ is typical for crustal rocks with fault normal stresses
above 200 MPa, but lower than $\mu = 0.85$ for smaller $\sigma_n$
\citep{byerlee1978}, but may be appropriate for an immature crustal
fault.

@@

However, because $M_{zz}$ is greater than $M_{H{\mathrm{max}}}$
underneath the higher topography, topographic stresses can only lead to
reverse faulting on the Wenchuan faults if subhorizontal shear zones or
channels are present, and are very weak so shear can occur at 
low $\tau$ relative to $\sigma_n$. This conclusion has been reached in
studies of generalized orogens \citep[e.g.,][]{fleschbendick2012} and
gravitational collapse of volcanic edifices \citep[e.g.,][]{byrne2013},
indicating that is a general requirement of gravitational spreading. Our
calculations of $\mu$ and $\phi$ are based on $T$, which is biased
towards the high slip patches in the upper crust (Equation
\ref{eqn:rake_misfit}).  The resulting low values of $\mu$ and low
to moderate values of $\phi$, probably yield stress conditions insufficient
for slip at high pressures found at depths > 10 km.
Similarly, a low viscosity horizontal
channel that underlies eastern Tibet, likely much deeper than the
seismogenic zone we consider here, would be able to flow in response to
$\tau$ regardless of $\sigma_n$ and may be able to facilitate
gravitational spreading of the orogen \citep[e.g.,][]{clark2005, cook2008,
fleschbendick2012}.

@@

Assuming that the Beichuan faults were at a Mohr-Coulomb fault criterion
immediately prior to the Wenchuan earthquake, we estimate the tectonic
stresses required for the faults to fail. We use a Bayesian estimation,
resulting in posteriors representing likelihood of tectonic stress,
static friction, and a pore pressure parameter. The posteriors
indicate that the maximum tectonic stress is oriented
$\sim$E-W and has a likely minimum of about 10 MPa per kilometer of depth
(i.e. $T^\prime_{\mathrm{max}}$ of at least 0.4).
The minimum tectonic stress is oriented $\sim$N-S and is
fairly low, with the most likely values lower than 10-12 MPa per kilometer of
depth ($T^\prime_{\mathrm{min}} <$ 0.5).
The highest likelihood coefficient of static friction on the fault is
estimated at about 0.2--0.3, although values up to 0.5--0.6 are
permissible. Fluid pressures are likely 0--0.5 of the total pressure.
Slip occurred on these faults instead of more favorably-oriented faults
elsewhere in the region, due to the inferred low coefficient of friction
and moderate fluid pressures.

@@

\section{Green's functions for point-source
loads}\label{greens-functions-for-point-source-loads}
For completeness, we reproduce the Boussinesq \citep[e.g.,][]{jeffreys1970}
and Cerruti \citep[e.g.,][]{love1927} solutions here.
Note that in these solutions, $\lambda$ and $\mu$ are re-defined, and
are the first and second Lame's parameters, respectively, instead of
rake and fault friction as in the body of the manuscript.

>>>>

% document
\begin{abstract}
Though it is widely recognized that large mountain ranges produce
significant stresses in the Earth's crust, these stresses are not
commonly quantified.  Nonetheless, near large mountains topography may
affect fault activity by changing the stress balance on the faults. In
this work, we calculate the stress field from topography in the
Longmen Shan (Sichuan, China) and resolve those stresses on several
models of the faults that ruptured in the 2008 M$_\mathrm{w}$ 7.9
Wenchuan earthquake. We find that the topography results in shear
stresses up to 20 MPa and normal stresses up to 80 MPa on the faults,
with significant variability across the faults. Topographic stresses
generally load the fault in a normal and left-lateral shear sense,
opposite to the inferred coseismic slip sense, and thus inhibit the
coseismic slip. We estimate the tectonic stress needed to overcome
topographic and lithostatic stresses by assuming that the direction of
maximum shear accumulated on the faults is roughly collinear with the
inferred coseismic slip.  We further estimate the static friction and
pore fluid pressure assuming that the fault was, on average, at
Mohr-Coulomb failure at the time of the Wenchuan earthquake. We use a
Bayesian inversion strategy, yielding posterior probability
distributions for the estimated parameters. We find most likely
estimates of maximum tectonic compressive stress near 0.6 $\rho g z$
and oriented $\sim$E-W, and minimum tectonic stress near 0.2 $\rho g
z$. Static friction on the fault is near 0.2, and pore fluid pressure
is between 0 and 0.4 of the total lithostatic
pressure.  \end{abstract}

@@

In areas of substantial relief, high terrain and steep slopes generate large
stresses in the crust beneath and adjacent to the high topography
\citep{jeffreys1924, coblentz1996}. Because of the irregularity of topography
in mountainous regions, the stresses produced by topography are also
heterogeneous, and may play a prominent role in local or regional deformation,
particularly if the region is tectonically active.  For example, shear and
normal stresses on a fault due to topographic loading may push a particular fault
closer to or farther from failure, or reorient the
net shear stress direction on a fault. These effects may affect the
localization or deformational style in a region. Furthermore, heterogeneous
topographic stresses on a particular fault may affect the way earthquake
ruptures propagate across the fault plane. Despite this, the degree to which
topographic stresses affect faulting has received little direct study.

In this work, we investigate the effects of topographic stresses on
the faults that ruptured in the 2008 \emph{M}7.9 Wenchuan, China
earthquake.  This earthquake is an ideal candidate for this study
because it occurred at the base of the Longmen Shan, one of the
largest and steepest escarpments on Earth (Figure~\ref{fig:lms_map}),
and has a well-studied coseismic slip distribution characterized by
significant along-strike variations in coseismic slip.  Additionally,
because the earthquake occurred after $\sim$2000 years of seismic
quiescence, postseismic stresses in the lithosphere are likely to be
negligible, suggesting that the stress state on the faults at the time
of rupture can be approximated as the sum of topographic, lithostatic,
and accumulated tectonic stresses. We then use assumptions that
coseismic slip is collinear with the direction of total accumulated
shear stress on the fault and that the fault is on average at a
Mohr-Coulomb failure criterion in order to bracket the tectonic stress
field, pore fluid pressure, and static friction of the fault, so that
these parameters are consistent with the toopographic stresses and
coseismic slip on the Wenchuan earthquake faults.

@@

Topographic stresses are only one component of the total stress field
in the crust \citep{molnar1988}. Coseismic slip in the Wenchaun
earthquake is due to the total accumulated stress on the faults, which
also includes components from lithostatic and tectonic stresses. (In
the present study, we do not consider stresses due to flexure
\citep[e.g.,][] {luttrell2007} or buoyancy
\citep[e.g.,][]{luttrell2011}.) By quantifying both the topographic
and lithostatic stresses, we can use coseismic slip models to solve
for the tectonic stress assuming that (1) coseismic slip is in the
direction of the the total shear stress on the fault \citep[e.g.,]
[]{angelier1994} and (2) the fault is at Mohr-Coulomb failure
everywhere that it slipped.  Assumption (1) is the common
`Wallace-Bott' assumption (named after \citet{wallace1951} and
\citet{bott1959}), and note that it ignores dynamic stresses during
the earthquake process, and only consider that fault slip is collinear
with the accumulated stress on the fault prior to rupture. While this
assumption should be subject to further testing, it is standard in all
studies that infer stress from earthquake data
\citep[e.g.,][]{mckenzie1969, angelier1994, michael1987, reches1987,
  luttrell2011, medinaluna2013}.  If topographic stresses are
significant and produce shear in the direction of fault slip, then for
given values of static friction and pore fluid pressure, we can
calculate the amount of tectonic stress that can be added to the
ambient stress field before the faults should rupture; given limited
acceptable ranges for friction and fluid pressure, we are essentially
able to place maximum constraints on tectonic stress. Alternately, if
topographic stresses work against coseismic slip, for given friction
and fluid pressures we can estimate the minimum magnitudes of tectonic
stresses necessary to overcome shear and frictional resistance to
slip. In a scenario with complex faulting and topography, it may be
possible to put bounds on both minimum and maximum magnitudes, in
addition to the directions of tectonic stresses. To account for the
non-uniqueness of the solution, we use a sampling-based Bayesian
Monte-Carlo methodology to estimate posterior probability density
functions (PDFs) of tectonic stresses, static fault friction, and pore
fluid pressure.

@@

\begin{equation}
M(x, y, z) = G(x, y, z) * F(x, y) \; ,
\end{equation}
where $G(x,y,z)$ is a set of Green's functions for the six stress
tensor elements, and $F(x,y)$ is a topographic loading function,
described below. We assume compressive stresses are positive, and
$x>0$ is east, $y>$ is north, $z=0$ is mean sea level, and $z>0$ is
depth. \citet{liuzoback1992} show that $M(x,y,z)$ can be decomposed
into two components as

@@

\begin{equation}
M^C(x, y, z) = G_x^C(x,y,z) * F_{h, x}(x, y) + G_y^C(x,y,z) * F_{h, y}(x, y) \; ,
\end{equation}
where $G_i^C(x,y,z)$ are the Cerruti solutions for a horizontal point
source load in the $i$ direction on the halfspace surface (see
Appendix A1). The horizontal loading functions derived by
\citet{liuzoback1992} are given by

@@

Topographic stresses on the Wenchuan faults are calculated on point sets
representing the faults taken from coseismic slip models. We use six models,
those of \citet{shen2009}, \citet{feng2010}, \citet{zhang2011},
\citet{fielding2013} and two from \citet{qi2011}.
All of these models rely on geodetic data to some degree, but they do not all 
use the same data in their inversion (e.g., \citep{feng2010} uses a different
InSAR catalogue as the others). All use different inversion strategies, and
different fault geometries (the two models of \citep{qi2011} share a common
fault geometry but use different regularization in the inference of the slip
distribution, and we consider both their `rough' and `smooth' models here). By using a suite of models in our calculations, we can infer
that results which are persistent in most or all of the models are more robust,
while other results specific to only one of the coseismic slip models can be
more confidently linked to specifics of the model geometry or inferred slip
distribution in that model.

@@

In our stress calculations, we discard
points above 851 m below sea level, as this is above the depth at which
we compute $G^C(x,y,z)$. The six stress tensor components calculated at
the regular grid points are linearly interpolated to points describing
the faults. Because the fault points are completely surrounded by the
grid nodes at which topographic stresses were calculated and those nodes
are spaced \textless{}1 km apart, the fault points cannot be more than a
few hundred meters from the nearest grid node, so a higher order
interpolation is not necessary. We then project the topographic stress
tensor to fault normal stress, $\sigma_n^M$, down-dip shear stress,
$\tau_d^M$, and strike-slip shear stress, $\tau_s^M$, at each point in
describing the fault geometry (the superscripts on stress
symbols denote the origin of the stresses).

@@

\begin{equation} 
\tau = \mu ( \sigma_n - \sigma_p ) \; ,
\label{eqn:amonton_raw} 
\end{equation}
where $\mu$ is the coefficient of static friction on the fault and
$\sigma_p$ is the pore fluid pressure \citep[e.g.,][]{sibson1985}, and
assuming that cohesion is negligible. We
describe the pore fluid pressure using a scalar, $0 \leq \phi \leq 1$,
which is the pore fluid pressure as a fraction of total pressure, and so
the failure criterion is

@@

We estimate the tectonic stress tensor field, $\mu$, and $\phi$
consistent with published coseismic slip models of the Wenchuan
earthquake using a Bayesian estimation, resulting in samples of posterior probability 
density functions of the model parameters. 
We first estimate posteriors of the tectonic stress tensor field 
$T$ consistent with
the coseismic slip models, and then estimate $\mu$ and $\phi$ consistent
with Mohr-Coulomb failure. The nature of Bayesian estimation allows us
to quantify both the relative likelihoods of model parameters and the
tradeoffs between them.

@@

We invert topographic stresses and coseismic slip models for tectonics
stresses using Bayesian methods, and making the common `Wallace-Bott'
assumption (named after \citet{wallace1951} and \citet{bott1959}) that
slip on the fault occurs in the general direction of the maximum
resolved shear stress on the fault prior to the initiation of the
earthquake \citep[e.g.,][] {mckenzie1969, angelier1994}. We estimate
the tectonic stresses in light of the topographic stresses and slip
distributions through the relation

@@

Previous workers have demonstrated that by quantifying topographic
stress, other components in the Coulomb stress balance may be bracketed
\citep[e.g.,][]{cattin1997, lamb2006, luttrell2011}. Each of these studies
uses somewhat different approaches. Our approach is most similar to that
of \citet{luttrell2011}, although there are significant
differences: (1) We use the topographic stresss calculations proposed by
\citet{liuzoback1992}, whereas 
\citet{luttrell2011} only uses the vertical loading from topography,
equivalent to $M^B$ in equation (\ref{eqn:bous}). (2) We do not consider
bouyancy forces due to lateral variations in density, due for instance to
Moho variation, as done by \citet{luttrell2011}. In the
Longmen Shan region, Moho variation is small compared to the change in
Moho depths at the Andean plate boundary, and so the buoyancy terms
should be relatively small. (3) We consider a full range of tectonic
stresses, instead of simply calculating the minimum principal tectonic
stress and its orientation. (4) We consider the stress tensor at each
point due to topographic loading, lithostatic stress, and horizontal
tectonic stress, and use inferred coseismic slip rake as a constraint on
the allowable stresses, rather than inferring the earthquake stress drop
from the coseismic slip models, as done by 
\citet{luttrell2011}. (5) We use both normal and shear stresses to
constrain pore fluid pressure and friction. (6) We use a Bayesian
estimation, resulting in samples of a PDF of tectonic stress, as well as $\mu$ and
$\phi$, rather than just solving for the minimum tectonic stress
required for faulting, as done by \citet{luttrell2011}.

@@

The maximum tectonic stress, $T_{\mathrm{max}}$, is consistently
oriented roughly E-W in our results. This orientation is oblique to the
Longmen Shan, which produces oblique (right-lateral and reverse sense)
shear on the Beichuan fault. $T_{\mathrm{min}}$ is $\sim$N-S
oriented, and is a small fraction of lithostatic pressure. This
stress configuration is in close agreement with pre-earthquake stress
orientation
measurements near the rupture zone (Figure \ref{fig:lms_stress_map}),
mostly from borehole breakout data from 2-5 km depth
\citep{heidbach2009}, which is the zone of maximum slip in the coseismic
slip models. It is somewhat discrepant with stress orientations
estimated at $\sim$800 m depth adjacent to the Beichuan fault
in the WFSD-1 borehole of the Wenchuan Earthquake Fault Scientific
Drilling Project several years after the 2008 earthquake \citep{cui2014},
which show $\sigma_{H{\mathrm{\max}}}$ to be more orthogonal to the fault
trace, suggesting that much of the right-lateral component of shear
stress was released during the earthquake. Our results are also similar to
the orientations of total stress obtained by 
\citet{medinaluna2013}, although they were unable to constrain the
magnitudes of stresses.

@@

The magnitudes of $T_{\mathrm{max}}$ and $T_{\mathrm{min}}$ are
dominantly constrained on the low end by our analysis, which is apparent
by the sharp decrease in the frequency of $p(T_{\mathrm{max}}|D)$ below
about $T^\prime_{\mathrm{max}} = 0.5$ (Figure~\ref{fig:joint_posteriors}). 
These results indicate that $T_{\mathrm{max}}$ of at
least $\sim$13.25 MPa km$^{-1}$ is necessary to overcome
topographic stresses resisting reverse and right-lateral slip on the
faults. We find that the highest likely ratio of strike-slip to dip-slip
shear along the Wenchuan earthquake faults is close to 1. This is
similar to the inferences of strain accumulation rates inferred from
`squishy' block modeling by \citet{loveless2011}, who
concluded that the rate of slip deficit accumulation in the thrust and
dextral senses were approximately equal. It should be noted that the
tectonic stresses we estimate here represent the accumulated stresses
prior to the Wenchuan earthquake, and the relation of these stresses to any
accumulated slip deficit needs to be through a model of strain
accumulation.

@@

CFR indicates whether a fault should fail under a given stress state:
CFR \textgreater{} 1 indicates failure, while a CFR \textless{} 1
indicates fault stability. We then calculate the CFR on each point in
the model of the Beichuan fault (594 points describe the fault model of
\citet{zhang2011}) based on the full stress field $S$ at
each point on the fault, for each of the 1000 samples of $T$, $\phi$ and
$\mu$ drawn randomly from the posteriors. We call this CFR$_f$. Then,
using the same $S$ and $\phi$, we calculate the CFR on an optimally
oriented fault with $\mu=0.6$ and no cohesion, which we call CFR$_o$.
The orientation of the optimally oriented fault is determined as being
the angle $\beta$ away from $\sigma^S_1$, where $\beta = (\tan^{-1} \mu) / 2$
and is in the $\sigma_1$--$\sigma_3$ plane \citep[e.g.,][]{sibson1985}.
Note that $\mu=0.6$ is typical for crustal rocks with fault normal stresses
above 200 MPa, but lower than $\mu = 0.85$ for smaller $\sigma_n$
\citep{byerlee1978}, but may be appropriate for an immature crustal
fault.

@@

However, because $M_{zz}$ is greater than $M_{H{\mathrm{max}}}$
underneath the higher topography, topographic stresses can only lead to
reverse faulting on the Wenchuan faults if subhorizontal shear zones or
channels are present, and are very weak so shear can occur at 
low $\tau$ relative to $\sigma_n$. This conclusion has been reached in
studies of generalized orogens \citep[e.g.,][]{fleschbendick2012} and
gravitational collapse of volcanic edifices \citep[e.g.,][]{byrne2013},
indicating that is a general requirement of gravitational spreading. Our
calculations of $\mu$ and $\phi$ are based on $T$, which is biased
towards the high slip patches in the upper crust (Equation
\ref{eqn:rake_misfit}).  The resulting low values of $\mu$ and low
to moderate values of $\phi$, probably yield stress conditions insufficient
for slip at high pressures found at depths $>$ 10 km.
Similarly, a low viscosity horizontal
channel that underlies eastern Tibet, likely much deeper than the
seismogenic zone we consider here, would be able to flow in response to
$\tau$ regardless of $\sigma_n$ and may be able to facilitate
gravitational spreading of the orogen \citep[e.g.,][]{clark2005, cook2008,
fleschbendick2012}.

@@

Assuming that the Beichuan faults were at a Mohr-Coulomb fault
criterion immediately prior to the Wenchuan earthquake, we estimate
the tectonic stresses required for the faults to fail. We use a
Bayesian estimation, resulting in samples of posterior probability
distributions representing likelihood of tectonic stress, static
friction, and a pore pressure parameter. The posteriors indicate that
the maximum tectonic stress is oriented $\sim$E-W and has a likely
minimum of about 10 MPa per kilometer of depth
(i.e. $T^\prime_{\mathrm{max}}$ of at least 0.4).  The minimum
tectonic stress is oriented $\sim$N-S and is fairly low, with the most
likely values lower than 10-12 MPa per kilometer of depth
($T^\prime_{\mathrm{min}} <$ 0.5).  The highest likelihood coefficient
of static friction on the fault is estimated at about 0.2--0.3,
although values up to 0.5--0.6 are permissible. Fluid pressures are
likely 0--0.5 of the total pressure.  Slip occurred on these faults
instead of more favorably-oriented faults elsewhere in the region, due
to the inferred low coefficient of friction and moderate fluid
pressures.

@@

\section{Green's functions for point-source
loads}\label{greens-functions-for-point-source-loads} For
completeness, we reproduce the Boussinesq
\citep[e.g.,][]{jeffreys1970} and Cerruti \citep[e.g.,][]{love1927}
solutions here.  Note that in these solutions, $\lambda$ and $\mu$ are
the first and second Lame's parameters, respectively, instead of rake
and fault friction as in the body of the manuscript.
